Quantifying unmet ICU beds in tertiary hospitals in urban India using Prospective Score Matching 

### 2022-09-11 manuscript.Rmd
sparar en github kopia. ganska babbligt ännu.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

@@

In order to improve intra-hospital care, trauma quality improvement programs (TQIP) have been developed. The aim of TQIPs is to ensure equivalent care and improve trauma care by developing best practice guidelines@tqip. TQIPs utilizes a retrospective cohort study, analysing the current of trauma care patients and post-fact attempting to determine the appropriate level of care for each patient. From these analyses many aspects of trauma care can be improved and optimized, such as determining the appropriate amount of ICU beds and other trauma care resources.@tqip

The existing TQIP often originate in high-income countries, and are either not applicable nor easily transferred to the hospital conditions of low and middle income countries (LMIC). This study aims to make an TQIP assessment in an Indian context and fill part of that knowledge gap, and to test whether there is merit in the testimony of the Indian colleagues. if there is indeed a shortage of ICU beds in Indian hospitals, further quantify the scope of that shortage. If this analysis indeed indicate a shortage, it can be used to motivate entending ICU resources and reduce preventable trauma morbidity and mortality.

@@

<!-- Present key elements of study design -->

@@

<!-- Describe the setting, locations, and relevant dates, including
periods of recruitment, exposure, follow-up, and data collection -->

@@

<!-- Cohort study: Give the eligibility criteria, and the sources and
methods of selection of participants. Describe methods of
follow-up. For matched studies, give matching criteria and number of
exposed and unexposed -->

@@

@@

<!-- Describe any efforts to address potential sources of bias -->

@@

<!-- Explain how the study size was arrived at -->

@@

<!-- Explain how quantitative variables were handled in the
analyses. If applicable, describe which groupings were chosen and why
-->

@@

(a) Describe all statistical methods, including those used to control
for confounding

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
table_nums <- captioner::captioner(prefix = "Tab.")
```

@@

In order to improve intra-hospital care, trauma quality improvement programs (TQIP) have been developed. The aim of TQIPs is to ensure equivalent care and improve trauma care by developing best practice guidelines@tqip. TQIPs utilizes a retrospective cohort study, analysing the current of trauma care patients and post-fact attempting to determine the appropriate level of care for each patient. From these analyses many aspects of trauma care can be improved and optimized, such as determining the appropriate amount of ICU beds and other trauma care resources @tqip.

Existing TQIP often originate in high-income countries, and are either not applicable nor easily transferred to the hospital conditions of low and middle income countries (LMIC). This study aims to make an TQIP assessment in an Indian context to fill part of that knowledge gap, and to test whether there is merit in the testimony of the Indian colleagues. If there is indeed a shortage of ICU beds in Indian hospitals, the study shall further quantify and propose a scope of the unmet need for ICU beds for trauma patients at tertiary hospitals in urban India. If this analysis indeed indicate a shortage, it can be used to motivate extending ICU resources and reduce preventable trauma morbidity and mortality.

@@

<!-- Present key elements of study design -->
Observational retrospective cohort study from data collected at four tertiary hospitals in urban India.

@@

<!-- Describe the setting, locations, and relevant dates, including
periods of recruitment, exposure, follow-up, and data collection -->
FRÅGA: kopierat rakt av härifrån https://github.com/titco/titco-I
Men det är ju exakt vad som ska under den här rubriken. Får man göra så eller är det plagiat?
The TITCO-I v1 dataset includes data on 16 000 patients admitted to four public university hospitals in urban India. The original dataset included 16 047 patients but 47 patients were deleted from the dataset for anonymization purposes, see the anonymization section for more details. At the time of the research Jai Prakash Narayan Apex Trauma Center (JPNATC), All India Institute of Medical Sciences, New Delhi, was a dedicated trauma centre with almost 180 beds. King Edward Memorial hospital (KEM), Mumbai, was a tertiary level hospital with no dedicated trauma ward. Lokmanya Tilak Municipal General Hospital (LTMGH), Mumbai was a tertiary level public university hospital with a dedicated trauma ward with 14 beds. The Institute of Post-Graduate Medical Education and Research and Seth Sukhlal Karnani Memorial Hospital (SSKM), Kolkata was a tertiary level public university hospital, but without a dedicated trauma ward. The cost of care in these hospitals was nominal and the patients admitted represented mainly a lower socioeconomic stratum of the population. The first patient was enrolled in July 2013 and the last patient in December 2015. The exact enrollment period varied across participating hospitals.

@@

<!-- Cohort study: Give the eligibility criteria, and the sources and
methods of selection of participants. Describe methods of
follow-up. For matched studies, give matching criteria and number of
exposed and unexposed -->
The TITCO1 data set includes all data recorded for trauma patients at the participating hospitals is included in the TITCO1 study.

@@

There are 193 possible observations recorded for each patient. A general presentation of these observations is in table_nums('tab_1')

tab.1_cap <- table_nums(name = "tab_1", 
                        caption = "Data types collected in the TITCO1 dataset")

                     knitr::kable(table1_1, "simple")

datatvätt:
FRÅGA: quantitativa variablerna har många NA som jag tror egentligen är NO, ex om pat har haft syrgas vid spO2 mätning. kan jag tänka lite fritt här när det känns uppenbart eller ska jag förankra? också finns det ex obs där numeriska värdet för spO2 saknas men ändå data yes/no om pat fått syrgas och jag vet inte om sån data är meningsfull?
FRÅGA: vad betyder GCS verbal V1?  det är inte felregistrerat det är ett alternativ i codebook. jag har hittat ett förslag om att det är pat som inte kan ge verbalt svar pga intub eller surgical airway, men det finns pat som har varken eller och GCS verbalt = 1v

The quantitative data include recorded parameters such as vital parameters, lab results, GCS total score and cumulative time in surgery or ICU, if applicable. Measurement or assessment techniques are standardized.

The qualitative data include categorical data such as coarse categorization or the accident mechanism and injury, details about arrival circumstances, and whether or not a patient had received a medical examination or intervention.

Text data are freehand examination results and operation categorizations, as well as assigned ICD codes. These data are vast and contain a lot of useful information, but also introduce a lot of challenges. There is a lack of standardization, user dependence, heterogenous grouping caused by using too general ICD codes, and datamining challenges for this type of data. If it may cause inaccuracies or practical challenges, it shall be omitted.

The intermediate outcome of interest is ICU admission, irregardless of duration, and to identify the predictors that warrant ICU admission. Therefore, the properties of the patients that were admitted into the ICU will be examined. The variables with the highest propensity to warrant a ICU admission to be ICU - predictors - will be identified. 
Using these predictors and propensity score matching will identify patients that weren’t admitted into the ICU, but presented with a set of predictors - a propensity score - very similar to those that were admitted. The ultimate outcome of this analysis is to quantify this group of patients and consequently quantify the unmet need for ICU beds.

@@

<!-- Describe any efforts to address potential sources of bias -->
The TITCO1 dataset are observational data collected from a clinical environment and not an experimental setting. Producing a randomized control trial study for these types of lifesaving medical interventions for trauma patients would obviously be morally indefensible. Medical interventions, examinations and ICU admission are not randomized, they are a consequence of a patients medical condition. There exists a causal effect between the independent and the dependent variables: a patient is sent to the ICU precisely because of the severity of their condition.

This causality introduces biases that must be addressed in the statistical analysis. Traditionally, multivariable regression has been used tocontrol for confounding in these cases but methods based on propensity scores have become popular alternatives. Propensity Score Matching will be descibed in detail in chapter Statistical methods.

@@

<!-- Explain how the study size was arrived at -->
The data of 16047 trauma patients were collected at the participating hospitals over 2.5 years, as described in chapter Settings
47 entries have been randomly excluded for anonymization purposes, and for the purpose of this study the 120 patients who were dead on arrival have been excluded aswell.

@@

<!-- Explain how quantitative variables were handled in the
analyses. If applicable, describe which groupings were chosen and why
-->
Fråga: hur diffar detta från vad som finns i Variables and data sources/measurements

@@

(a) Describe all statistical methods, including those used to control
for confounding
-->
Propensity score matching will be used in this analysis. Propensity score matching (PSM) is typically performed to determine the efficacy of a conditionalized intervention - causality between the dependent variables and available independent variables - where there is a need to offset or balance the effect of this causality. 
<!-- 
The goal of propensity score matching is usually to identify suitable control subjects that did not receive an intervention, even though they fairly well met the same inclusion criteria that the subjects that did receive the intervention had. Control subjects who are similar enough to the intervention group in the most relevant aspects for receiving treatment. 
-->
In this context PSM could provide tools for an unbiased examination of the efficacy of ICU care on mortality. When simply comparing the cohort that did go to the ICU to the cohort that did not, it would appear that ICU care actually increased mortality. In the TITCO1 dataset the survival rate of patients who did not go to the ICU was 81% versus 68% for patients who did. This is of course an incorrect conclusion, the higher mortality rates in the ICU is because patients who are admitted here are worse off. If, however, the conditions that typically warrant ICU admission could be identified, a subset of the TITCO1 dataset that fairly well met those criteria but did not go to the ICU could be identified using propensity score matching.

Identifying a subset of patients that should have received ICU care but didn't is the end point for this thesis.
In the typical PSM context this subset is further used as the control group to determine the efficacy of ICU care.

The first step is to identify predictor variables: independent variables most predictive of the outcome or dependent variable. In this context it means to first identify the conditions and medical interventions received that correlate to ICU care. All properties of the patients that were sent to the ICU will be compared to the patents who didn't, in order to identify properties that distinguishes this cohort.

Second step is to calculate propensity scores for all patients. Each patient in the TITCO1 data set gets a propensity score based on their medical conditions and interventions, specifically their individual values of the predictor variables.The propensity score determines the likelihood that a patient will receive ICU care, given their condition.

Third step is PSM. PSM identify members of the subset that did not get the intervention but had the same predictors as the subgroup that did get the intervention. The purpose of the propensity score matching is identifying patients that weren’t admitted into the ICU, but presented with a set of predictors - a propensity score - very similar to those that were admitted.
